---
title: "Reliability and Multi-Group CFA analysis"
author: "Anne-Flore Cabouat adapt Eliane"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)

# psych runs parallel analysis simulations on several cores on macOS/Linux,
# which makes the random results change from run to run even with set.seed().
# One core gives reproducible results, identical to those obtained on Windows.
options(mc.cores = 1)

# Paths are relative to the repository root (instrument.Rproj)
phase_dir  <- here("phase3-validation")
data_dir   <- file.path(phase_dir, "00_collapse-categories", "output")
output_dir <- file.path(phase_dir, "05_cfa-multigroup", "output")

# File names use system names in snake_case, e.g. "Power BI" -> "power_bi"
slug <- function(x) gsub("[^a-z0-9]+", "_", tolower(x))

if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
}

```

## Librairies

```{r librairies}
library(dplyr)
library(knitr) #RStudio
library(png) #output png
library(corrplot) #plot correlation matrix
library(misty) #Little's test for MCAR
#library(mice)
library(mifa) # MICE for missing data handling
library(psych) # for EFA
library(RColorBrewer) #for corr matrix
library(lavaan) # for cfa
```
## Functions
```{r functions}


make_corrM <- function(d = NULL, order = NULL, ...) {
 
  #if (is.null(d)) {
  #  if (exists("r_mat")) return(r_mat)
  #  return(NULL)
  #}
  
  
  print("corr M linha 55")
  
   # Mantain the original column names 
  d <- as.data.frame(d)
  
  # 2. Remove items with 0 variance 
  valid_cols <- sapply(d, function(col) length(unique(na.omit(col))) > 1)
  d_clean <- d[, valid_cols, drop = FALSE]

  
  if (ncol(d_clean) < 2) return(NULL)
  
 
  
  # 3. Convert to orederd factors and mantain the column names 
  orig_names <- colnames(d_clean)
  d_ordered <- as.data.frame(lapply(d_clean, ordered), check.names = FALSE)
  colnames(d_ordered) <- orig_names
  
  
  
  
  # 4. Calculate the polychoric correlation with pairwise deletion
  het <- tryCatch({
    polycor::hetcor(
      d_ordered, 
      use = "pairwise.complete.obs", 
      std.err = FALSE, 
      ML = FALSE
    )
  }, error = function(e) NULL)
  

  
  
  if (is.null(het)) return(NULL)
  
  corr_matrix <- het$correlations
  
  # Original names in columns and rows of the matrix 
  dimnames(corr_matrix) <- list(orig_names, orig_names)
  
  # 5. Reordering opcional
  if (!is.null(order) && length(order) > 0) {
    valid_order <- intersect(order, colnames(corr_matrix))
    if (length(valid_order) > 1) {
      corr_matrix <- corr_matrix[valid_order, valid_order, drop = FALSE]
    }
  }
  
  return(corr_matrix)
}

# Applicability of EFA

## Bartlett test of sphericity - from TH
bartlettTest <- function(corrMatrix,data){
  bart <- cortest.bartlett(corrMatrix, n = nrow(data))
  if(bart[2]>0.05) cat("WARNING the p value is above 0.05") else cat("The p value is below 0.05. We are good to continue.")
  cat("\n\n")
  print(bart)
  return(bart)
}

##KMO to complement Bartlett - from TH
KMOTest <- function(corrMatrix){
  kmo <- KMO(corrMatrix)
  cat(paste("The overall measure of sampling adequacy is: ",kmo[1]))
  cat("\n\n")
  if(kmo[1]<.7) cat("WARNING the sampling adequacy has dropped below 0.7") else cat("The sampling adequacy is above 0.7. We are generally good.")
  cat("\n\n")
  return(kmo)
}

# Missing values

missingTest <- function(na_percent, mcar_pvalue){
  if (na_percent<5 & mcar_pvalue<=.05){
      cat("Data is MCAR and less than 5% missing data, which is negligible. We are good to continue.")
  } else if (na_percent<5 & mcar_pvalue>.05){
      cat("Data is NOT MCAR but less than 5% missing data, which is negligible. We are good to continue.")
  } else if(na_percent>=5 & mcar_pvalue>.05) {
      cat("WARNING missing data is not negligible (> 5%) AND missing data is NOT MCAR. Further investigation is needed.")
    }else if (na_percent>=5 & na_percent<40 & mcar_pvalue<=.05){
      cat("Data is MCAR but more than 5% missing data. Use imputation / likelihood methods.")
    }  else {
      cat("40% or more missing data and / or something is wrong.")
    }}


# Parallel analysis for each group

parallelAnalysis <- function(corrMatrix, observations = nrow(the_data), file_suffix){
  pdf(file.path(output_dir, paste0("scree_plot_", slug(file_suffix), ".pdf")), width=8, height=4)
  parallel <- fa.parallel(corrMatrix, n.obs=observations, fa="fa", n.iter=100, main="Scree plots with parallel analysis", error.bars = TRUE)
  print(parallel)
  dev.off() 
  cat("\n\n")
}

stimulus_data_test <- function(the_data, the_letter){
  ## Here wwe run test on the data
  ### Shapiro Wilk’s test (P-values < 0.05 indicate not normal)
  cat(paste("\n=================\n NORMALITY in", the_letter, "\n"))

  cat(paste("\n #### Shapiro in", the_letter, "\n"))
  mapply(shapiro.test, the_data)

  ### Multivariate normality
  # To say the data are multivariate normal: 
  # • z-kurtosis < 5 (Bentler, 2006) and the P-value should be ≥ 0.05.
  # • the plot should also form a straight line (Arifin, 2015).

  cat(paste("\n #### Multivariate normality in", the_letter, "\n"))
  mardia(the_data)

  cat(paste("\n=================\n MISSING DATA IN in", the_letter, "\n"))
  ### Missing data

  if (anyNA(the_data)) {
  cat("#### Little's test for MCAR\n")
  #test MCAR
  MCAR <- na.test(the_data)
  mcar_p <- MCAR[["result"]][["pval"]]
  
  cat("\n#### Amount of missing data\n")
  nb_obs <- nrow(the_data)*ncol(the_data)
  na_amount <- sum(is.na(the_data))/nb_obs*100
  cat(paste(as.numeric(na_amount)), "% of data is missing\n\n")
  
  missingTest(na_amount, mcar_p)
  } else {
  cat("No missing values found in the_data.\n")
  }


### Multiple Imputation for missing values
  cat(paste("\n Multiple Imputation for missing values in", the_letter, "\n"))
  
  print("4 corr M sem parametro")
  #corr_M <- make_corrM(the_data, nb_fa = 4) PREVIS
  corr_M <- make_corrM()
 

  cat(paste("\n=================\n FACTORABILITY in", the_letter, "\n"))
  
  ####  Bartlett’s test of sphericity
  # An objective test of the factorability of the correlation matrix is Bartlett’s (1954) test of sphericity, which statistically tests the hypothesis that the correlation matrix contains ones on the diagonal and zeros on the off-diagonals. Hence, that it was generated by random data. This test should produce a statistically significant chi-square value to justify the application of EFA.
  # If the p-value from Bartlett’s Test of Sphericity is lower than our chosen significance level (common choices are 0.10, 0.05, and 0.01), then our dataset is suitable for a data reduction technique.
  cat(paste("\n #### Bartlett in", the_letter, "\n"))
  bartlettTest(corr_M, the_data)

  ####  KMO
  # Large sample sizes make the Bartlett test sensitive to even trivial deviations from randomness, so its results should be supplemented with a measure of sampling adequacy. The Kaiser-Meyer-Olkin (KMO; Kaiser, 1974) measure of sampling adequacy is the ratio of correlations and partial correlations that reflects the extent to which correlations are a function of the variance shared across all variables rather than the variance shared by particular pairs of variables. KMO values range from 0.00 to 1.00 and can be computed for the total correlation matrix as well as for each measured variable.
  # - KMO values ≥.70 are desired
  # - KMO values ≤.50 are generally considered unacceptable
  cat(paste("\n #### KMO in", the_letter, "\n"))
  KMOTest(corr_M)


  
  parallelAnalysis(corr_M, observations=nb_obs, "All")

}

fitMeasures_output <- function(modelFit, model_name, file_label = slug(model_name)) {
  print(fitMeasures(modelFit, c("chisq", "df", "pvalue", "cfi", "rmsea"),
                  output = "text"), add.h0 = TRUE)

  fit_Indices <- fitMeasures(modelFit, output = "matrix")
  colnames(fit_Indices)[1] <- c(model_name)
  write.csv(fit_Indices, file = file.path(output_dir, paste0("fit_indices_", file_label, ".csv")))
  
  
  #file = file.path(pasta_saida, nome_arquivo)
  
  
  cat("\nDETAILS\n=================\n")
  print(fit_Indices)
  
  # cat("\n RMSEA - \n")
  # this_RMSEAIndices <- fit_Indices[c("rmsea","rmsea.pvalue","rmsea.ci.lower","rmsea.ci.upper","rmsea.ci.level")]
  # print(this_RMSEAIndices)
  # 
  # cat("\n SRMR \n")
  # this_SRMRIndices <- fit_Indices[c("srmr")]
  # print(this_SRMRIndices)
  # 
  # cat("\n Comparative Fit Index \n")
  # this_CFIIndices <- fit_Indices[c("cfi")]
  # print(this_CFIIndices)
  # 
  # cat("\n Chi-square \n")
  # this_ChiSquareIndices <- fit_Indices[c("chisq","df","pvalue")]
  # print(this_ChiSquareIndices)

}

select_all_variables <- function (dataset, col_list) {
  dataset %>% select(all_of(unlist(col_list)))
}

```

## Multigroup CFA

https://lavaan.ugent.be/tutorial/groups.html

```{r prepare model and data filters}

#we retrieve our model from EFA conducted on all data

this_model <- "Tested_model" #model used in the validation survey

#UX_factors <- list(
#  visualization = list('P106','P089','P098','P097', 'P022A','P067'),
#  cognitive = list('P064', 'P301', 'P060', 'P027'),
#  intuitiveness = list('P088', 'P026', 'P4026A')
#)


UX_factors <- list(
  visualization = list('VisFaithful', 'VisSimpleUnderstand', 'VisAppropriate', 'VisSysRetrieveInfo', 'VisRemember', 'VisTrialError'),
  cognitive = list('CogSysThinkDeep', 'CogSysUseful', 'CogSysEnjoy', 'CogTaskComplete'),
  intuitiveness = list('IntSysSimpleUnderstand', 'IntSysEasyStart', 'IntAnalysisConfigure')
)


# Create a vector of column names to keep for the model with selected items
columns_to_keep <- unlist(UX_factors)

# Function to convert a factor list to the required string format
factor_to_string <- function(factor_name, factor_list) {
  paste(factor_name, " =~ ", paste(factor_list, collapse = " + "))
}

```
## Load data

```{r load data}
stimuli_letters <- list("Tableau","JMP", "Excel", "Data Studio","Power BI")

grp_data <- read.csv(file.path(data_dir, "multigroup_ratings.csv"), fileEncoding = "UTF-8-BOM")
data <- subset(grp_data, select = -c(id, seed, stimulus))  # item answers only

# and make directories if needed



generated_data_dir  <- output_dir
generated_plots_dir <- output_dir

# Cria as pastas caso não existam
if (!dir.exists(generated_data_dir)) {
  dir.create(generated_data_dir, recursive = TRUE)
}

if (!dir.exists(generated_plots_dir)) {
  dir.create(generated_plots_dir, recursive = TRUE)
}



```

## Descriptive statistics - Check minimum/maximum values per item, and screen for missing values
```{r data overview}

cat("\n\n==========================================\n
    ==========================================\n
    ==========================================\n")
cat("\n\n OVERVIEW \n\n")
cat("==========================================\n
    ==========================================\n
    ==========================================\n\n")

describe(data)
response.frequencies(data)
nb_obs <- nrow(data)
cat("\n Sample size:")
print(nb_obs)

```


```{r reliability of each items (each factor)}

cat("\n\n==========================================\n")
cat("==========================================\n")
cat("==========================================\n")
cat("\n \n CORRELATIONS \n \n")
cat("==========================================\n")
cat("==========================================\n")
cat("==========================================\n\n")

colnames(columns_to_keep)

data_reduced_factors <- data %>% 
  dplyr::select(dplyr::all_of(unname(columns_to_keep)))


# Correlation Matrix with selected items
#data_reduced_factors <- data %>% dplyr::select(dplyr::all_of(columns_to_keep))

print("1 - antes make_corrM")

colnames(data_reduced_factors)

corrM <- make_corrM(data_reduced_factors)

## Save as CSV
#write.csv(corrM, "generatedData-CFA/subscales-correlation.csv")
write.csv(corrM, file.path(output_dir, "items_correlations.csv"))


## Save as PDF
this_file <- file.path(output_dir, "items-correlation-plot.pdf")
pdf(this_file, height = 15, width = 15)
plot <- corrplot(
  corrM,
  tl.srt      = 30,
  method      = "color",
  col         = COL2('RdYlBu'),
  order       = "original",
  addrect     = 4,
  tl.col      = "black", 
  tl.cex      = 1.3,
  addCoef.col = "white", 
  number.cex  = 1
)
dev.off()


cat("\n\n==========================================\n")
cat("==========================================\n")
cat("==========================================\n")
cat("\n \n RELIABILITY \n \n")
cat("==========================================\n")
cat("==========================================\n")
cat("==========================================\n\n")


# Protected reliability computation with zero-variance guards
make_reliability_df <- function(this_data, col_name, row_suffix = "") {
  raw_alpha_name <- paste0("raw_alpha", row_suffix)
  std_alpha_name <- paste0("std.alpha", row_suffix)
  tot_omega_name <- paste0("omega.tot", row_suffix)
  
  res_df <- data.frame(
    val = c(NA_real_, NA_real_, NA_real_),
    row.names = c(raw_alpha_name, std_alpha_name, tot_omega_name)
  )
  colnames(res_df) <- col_name
  
  # Remove constant items
  varying_cols <- sapply(this_data, function(x) length(unique(na.omit(x))) > 1)
  this_data_clean <- this_data[, varying_cols, drop = FALSE]
  
  if (ncol(this_data_clean) < 2 || nrow(na.omit(this_data_clean)) < 3) {
    return(res_df)
  }
  
  # 1. Cronbach's Alpha
  tryCatch({
    alpha_res <- suppressWarnings(
      psych::alpha(this_data_clean, check.keys = FALSE, warnings = FALSE)
    )
    res_df[raw_alpha_name, col_name] <- as.numeric(alpha_res$total$raw_alpha)
    res_df[std_alpha_name, col_name] <- as.numeric(alpha_res$total$std.alpha)
  }, error = function(e) {
    message("Alpha computation notice for ", col_name, row_suffix, ": ", e$message)
  })
  
  # 2. McDonald's Omega
  tryCatch({
    nb_obs <- nrow(this_data_clean)
    
    print("2 - antes make_corrM")

    colnames(data_reduced_factors)

    corrM_sub <- make_corrM(this_data_clean)
    
    if (!is.null(corrM_sub) && ncol(corrM_sub) >= 2) {
      if (any(eigen(corrM_sub)$values <= 0)) {
        corrM_sub <- psych::cor.smooth(corrM_sub)
      }
      
      om <- suppressWarnings(
        psych::omega(
          corrM_sub, 
          nfactors = 1, 
          n.obs = nb_obs, 
          plot = FALSE
        )
      )
      res_df[tot_omega_name, col_name] <- as.numeric(om$omega.tot)
    }
  }, error = function(e) {
    message("Omega computation notice for ", col_name, row_suffix, ": ", e$message)
  })
  
  return(res_df)
}

# Reliability data structure initialization
reliability_metrics <- c("raw_alpha", "std.alpha", "omega.tot")
subscale_reliability_df <- data.frame()

row_names <- as.vector(t(outer(reliability_metrics, stimuli_letters, function(m, l) paste(m, "in", l))))
subscale_group_reliability_df <- data.frame(
  matrix(NA_real_, nrow = length(row_names), ncol = length(names(UX_factors))),
  row.names = row_names
)
colnames(subscale_group_reliability_df) <- names(UX_factors)

# Compute reliability across factors and stimulus groups
for (factor in names(UX_factors)) {
  cat(paste("\n======================\nFactor reliability on selected items:", factor, "(Alpha & Omega)\n\n"))
  
  factor_items <- unlist(UX_factors[[factor]])
  factor_data  <- data %>% dplyr::select(dplyr::all_of(factor_items))
  
  factor_df <- make_reliability_df(factor_data, col_name = factor)
  if (ncol(subscale_reliability_df) == 0) {
    subscale_reliability_df <- factor_df
  } else {
    subscale_reliability_df <- cbind(subscale_reliability_df, factor_df)
  }
  
  for (letter in stimuli_letters) {
    the_data <- grp_data %>% dplyr::filter(stimulus == letter)
    factor_group_data <- the_data %>% dplyr::select(dplyr::all_of(factor_items))
    
    this_df <- make_reliability_df(
      factor_group_data,
      col_name   = factor,
      row_suffix = paste(" in", letter)
    )
    
    for (r in rownames(this_df)) {
      if (r %in% rownames(subscale_group_reliability_df)) {
        subscale_group_reliability_df[r, factor] <- this_df[r, factor]
      }
    }
  }
}

# Export consolidated metrics
reliability_df_full <- rbind(subscale_reliability_df, subscale_group_reliability_df)
write.csv(reliability_df_full, file.path(output_dir, "reliability-subscales.csv"))
```

```{r run CFA analysis}

cat("\n\n==========================================")
cat("\n==========================================\n")
cat("\n Model fitting on agg data across stimuli\n")
cat("==========================================\n")
cat("==========================================\n\n")

## Create the model from list UX_factors
# Apply the function to each element of the list
formatted_strings <- sapply(names(UX_factors), function(factor_name) {
  factor_to_string(factor_name, UX_factors[[factor_name]])
})

# Combine the strings into a single string
model_string <- paste(formatted_strings, collapse = '\n')

# Print the final string
cat(model_string)



# CFA General - without separating by system
fit_overall_pooled <- cfa(
  model   = model_string,
  data    = data,          # data contém todas as respostas juntas sem a coluna 'stimulus'
  missing = "fiml"
)

# Exibir e salvar os índices gerais
summary(fit_overall_pooled, fit.measures = TRUE, standardized = TRUE)
fitMeasures_output(fit_overall_pooled, paste0(this_model, "-POOLED_OVERALL"), "pooled")

# Filtra apenas os estímulos selecionados
grp_data_cfa <- grp_data %>%
  dplyr::filter(stimulus %in% stimuli_letters) %>%
  droplevels()

cat("\n Model fitting\n")
this_fit <- cfa(
  model   = model_string,
  data    = grp_data_cfa, 
  group   = "stimulus",
  missing = "fiml"
)
#
#  this_fit <- cfa(model_string,
#           data = grp_data, 
#           group = "stimulus",
#           missing = "fiml")

  
  
  
  
# get information per group
summary(this_fit)

cat("\n==============\n")
cat("Model fit indices")
cat("\n==============\n")

this_filename = paste(this_model, '-full_data', sep='')
fitMeasures_output(this_fit, this_filename, "multigroup")

  
```

### Model fit indices

### GOAL

In their reference work on cutoff criteria for fit indices, Hu and Bentler (1999) recommend a combination of two criteria to retain a model: fit indices such as TLI or CFI should be higher than .95, and SRMR should be lower than 0.9. They add that a combinational cutoff criterion of RMSEA at .06 and SRMR at .09 is possible but less desirable because it tends to reject models that are, in fact, good fits.
https://doi.org/10.1080/10705519909540118


#### SRMR (Standardized Root Mean Square Residual)
> srmr

SRMR is particularly useful for evaluating the discrepancy between the observed and predicted covariance matrices.

SRMR is computed by taking the square root of the mean squared differences between corresponding elements in these two matrices.

The SRMR is then standardized to be independent of the scale of the observed variables. This makes it more comparable across different datasets and models.

A value less than . 08 is generally considered a good fit (Hu & Bentler, 1999)

#### RMSEA (Root Mean Square Error of Approximation)
> rmsea, rmsea.pvalue, rmsea.ci.lowe,r rmsea.ci.upper, rmsea.ci.level

RMSEA is a measure of how well the model's implied relationships between variables match the observed relationships in the data. It measures the average magnitude of the differences between predicted and observed values. In the context of CFA, it looks at how well the model-predicted covariances or correlations match the actual observed covariances or correlations among the variables.

RMSEA values less than 0.05 are good, values between 0.05 and 0.08 are acceptable, values between 0.08 and 0.1 are marginal, and values greater than 0.1 are poor. https://psycnet.apa.org/doi/10.1037/1082-989X.4.3.272
Hu and Bentler (1999) however recommend CFI or TLI + SRMR as validation criterion.

#### CFI (Comparative Fit Index)
> cfi

CFI is calculated by comparing the fit of the hypothesized model to a null model that assumes complete independence among the observed variables. The index ranges from 0 to 1, with higher values indicating better fit. A common rule of thumb is that CFI values above 0.90 or 0.95 are indicative of an acceptable or good fit, respectively.

#### Chi-square
> chisq, df, pvalue

It is calculated by comparing the observed sample covariance matrix with the covariance matrix predicted by the specified model. The formula involves summing the squared differences between the observed and expected (model-implied) covariances, adjusted for the sample size.

The chi-square test requires specifying the degrees of freedom, which depend on the complexity of the model. A small chi-square value relative to its degrees of freedom suggests that the model fits the data well. Conversely, a large chi-square value may indicate a poor fit between the model and the data.

1. Significance Test: The chi-square statistic is associated with a p-value (pvalue) that tests the null hypothesis. If the p-value is small (typically below a chosen significance level, such as 0.05), it suggests that there is evidence to reject the null hypothesis, indicating that the model does not fit the data well. In your provided data, the p-value associated with chisq is 0.000, suggesting that the model significantly deviates from the observed data.

2. Model Complexity: The chi-square statistic is sensitive to model complexity, meaning that as the model becomes more complex (e.g., with more estimated parameters), the chi-square tends to increase. Therefore, larger chi-square values may be expected for more complex models.

3. Sample Size Influence: The chi-square test can be influenced by sample size. In large samples, even small discrepancies between the model and the data can lead to a significant chi-square. Researchers often consider this and use additional fit indices, such as the Comparative Fit Index (CFI), Root Mean Square Error of Approximation (RMSEA), and Standardized Root Mean Square Residual (SRMR), to assess model fit.


### Measurement invariance testing

https://lavaan.ugent.be/tutorial/groups.html#measurement-invariance-testing
NB: Intercepts represent the average score or level of the latent construct when all other variables (including other latent constructs and observed variables) are held constant.

Pr(>Chisq) tells us if the less constrained model fits the data significantly better (e.g., that fit1 fits the data better compared to fit2)

fit1 = basic model, per group analysis
fit2 = model with constraint on factor loadings (equal loadings among groups)
fit3 = model with constraint on both loadings and intercepts


```{r measurement invariance}
cat("\n====================\n")
cat("\n Measurement invariance\n")
cat("\n====================\n")
  
# 1. Garantir que a base contenha apenas os estímulos selecionados
grp_data_cfa <- grp_data %>%
  dplyr::filter(stimulus %in% stimuli_letters) %>%
  droplevels()

# 2. Configural Invariance (mesma estrutura fatorial em todos os grupos)
# Re-estimado na base filtrada para garantir consistência com fit2 e fit3
fit1 <- cfa(
  model   = model_string, 
  data    = grp_data_cfa, 
  missing = "fiml", 
  group   = "stimulus"
)

# 3. Metric / Weak Invariance (cargas fatoriais fixadas como iguais)
fit2 <- cfa(
  model       = model_string, 
  data        = grp_data_cfa, 
  missing     = "fiml", 
  group       = "stimulus",
  group.equal = "loadings"
)

# 4. Scalar / Strong Invariance (cargas e interceptos fixados como iguais)
fit3 <- cfa(
  model       = model_string, 
  data        = grp_data_cfa, 
  missing     = "fiml", 
  group       = "stimulus",
  group.equal = c("intercepts", "loadings")
)

# 5. Testes de Comparação Hierárquica de Modelos (Likelihood Ratio Test)
cat("\n--- Likelihood Ratio Tests (Model Comparison) ---\n")
lavTestLRT(fit1, fit2, fit3)
```

#We already ran checks on overall data. Now we also check individual subgroups
```{r individual group data normality and EFA appropriateness checks}

row_names <- c("alpha", "omega.tot (1 factor)", "omega.tot (3 factors)")
reliab_df <- data.frame(row.names = row_names)

nbs_factors <- c(1, 3)

for (letter in stimuli_letters) {
  cat(paste("\n=========================================="))
  cat(paste("\n Processing stimulus:", letter))
  cat(paste("\n==========================================\n"))
  
  the_data <- grp_data %>% 
    dplyr::filter(stimulus == letter) %>% 
    droplevels()
  
  if (nrow(the_data) == 0) {
    message("No data for stimulus: ", letter)
    next
  }
  
  # Select variables
  subset_data <- select_all_variables(the_data, columns_to_keep)
  
  # 1. Checagens normality and adequacy
  tryCatch({
    stimulus_data_test(subset_data, letter)
  }, error = function(e) {
    message("Warnign for preliminary tests ", letter, ": ", e$message)
  })
  

  this_alpha <- NA_real_
  this_om_1f <- NA_real_
  this_om_4f <- NA_real_
  
  for (nb_f in nbs_factors) {
    cat(paste("\n--- RELIABILITY with", nb_f, "factor(s) in", letter, "---\n"))
    
    tryCatch({
      
       
      colnames(subset_data)
      
      c_mat <- make_corrM(subset_data)
      
      #save correlation matrix:
      write.csv(c_mat, file.path(output_dir, paste0("corr_matrix_", slug(letter), ".csv")))

      
      if (!is.null(c_mat) && ncol(c_mat) >= 2) {
        # smoothing of non positive matrix
        if (any(eigen(c_mat)$values <= 0)) {
          c_mat <- psych::cor.smooth(c_mat)
        }
        
        om <- suppressWarnings(
          psych::omega(
            c_mat, 
            nfactors = nb_f, 
            n.obs    = nrow(subset_data), 
            rotate   = "Promax", 
            plot     = FALSE
          )
        )
        
        cat("Alpha:", om$alpha, "\n")
        cat("G.6:", om$G6, "\n")
        cat("Omega Total:", om$omega.tot, "\n")
        
        if (nb_f == 1) {
          this_alpha <- om$alpha
          this_om_1f <- om$omega.tot
        } else {
          this_om_4f <- om$omega.tot
        }
      }
    }, error = function(e) {
      message("Error to calculate Omega (", nb_f, " factors) in ", letter, ": ", e$message)
    })
  }
  
  # Store reliability data
  reliab_df[[paste(letter, "- factors")]] <- c(this_alpha, this_om_1f, this_om_4f)
  
  # 3. CFA individual by stimulus
  cat(paste("\n--- Individual CFA Fit for", letter, "---\n"))
  tryCatch({
    this_fit <- lavaan::cfa(
      model   = model_string,
      data    = the_data, 
      missing = "fiml"
    )
    
    this_filename <- paste0(this_model, '-', letter)
    fitMeasures_output(this_fit, this_filename, slug(letter))
  }, error = function(e) {
    message("Error in CFA for stimulus ", letter, ": ", e$message)
  })
}

# Save reliability table by stimulus
write.csv(reliab_df, file.path(output_dir, "reliability-global-by-stimuli.csv"))

```


```{r omega sem, eval=FALSE}

# Bifactor / Hierarchical Model: 1 General Factor (overall_UX) + Specific Subfactors

# 1. Define the general latent factor syntax (valid identifier without spaces)
overall_items <- paste(unlist(UX_factors), collapse = " + ")
overall_factor_string <- paste0("overall_UX =~ ", overall_items)

# 2. Define specific factor syntax for each subscale
factor_strings <- sapply(names(UX_factors), function(factor_name) {
  paste0(factor_name, " =~ ", paste(unlist(UX_factors[[factor_name]]), collapse = " + "))
})

# 3. Combine factor syntax into a unified lavaan model specification
global_model_string <- paste(c(overall_factor_string, factor_strings), collapse = "\n")

cat("--- Bifactor Model Specification ---\n")
cat(global_model_string, "\n\n")

# 4. Filter dataset to include only valid stimulus groups and drop unused levels
grp_data_cfa <- grp_data %>%
  dplyr::filter(stimulus %in% stimuli_letters) %>%
  droplevels()

# 5. Fit the CFA model on pooled data across selected stimuli
# orthogonal = TRUE ensures general and specific factors are uncorrelated (standard bifactor assumption)
global_fit <- lavaan::cfa(
  model      = global_model_string, 
  data       = grp_data_cfa, 
  missing    = "fiml",
  orthogonal = TRUE
)

# Display model summary with standardized loadings and fit measures
summary(global_fit, fit.measures = TRUE, standardized = TRUE)

# 6. Estimate McDonald's Omega coefficients from the SEM object
cat("\n--- Omega from SEM Results ---\n")
om_sem <- psych::omegaFromSem(
  global_fit, 
  flip = TRUE, 
  plot = TRUE
)

print(om_sem)

```
