---
title: "Polychoric correlations — SUS, UEQ, Cognitive value, Intuitiveness, Visualization"
output: html_document
params:
  scales: "uxvis"   # "uxvis": the 13 UXVis items; "all": UXVis + SUS + UEQ-S items
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE)
# for data analysis
library(polycor)   # hetcor() for polychoric correlations
library(dplyr)
library(tidyr)

# for plots
library(ggplot2)

# to knit
library(knitr)
library(here)

stopifnot(params$scales %in% c("uxvis", "all"))

# Paths are relative to the repository root (instrument.Rproj)
phase_dir  <- here("phase3-validation")
input_file <- file.path(phase_dir, "00_collapse-categories", "output", "uxColaps_sus_ueqs_ratings.csv")
output_dir <- file.path(phase_dir, "03_polychoric-correlations", "output",
                        paste0(params$scales, "-scales"))
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
```

## Item groups

- **SUS**: `SUS_1`–`SUS_10`
- **UEQ_pragmatic**: `UEQ_1`–`UEQ_4`
- **UEQ_hedonic**: `UEQ_5`–`UEQ_8`
- **Cognitive_value**: `P064`, `P301`, `P060`, `P027`
- **Intuitiveness**: `P088`, `P026`, `P4026A`
- **Visualization**: `P106`, `P089`, `P098`, `P097`, `P022A`, `P067`

The complete polychoric matrix is calculated item by item (including, in this instance, pairs within the same construct), organized by construct blocks in the previously indicated order. Subsequently, a hierarchical cluster analysis is performed on this same matrix to verify whether the items group together empirically as hypothesized by the six intended constructs or if the data suggest a different structure. This notebook is run twice, set by the `scales` parameter: `uxvis` uses only the items of the UXVis scale (this run is used by step 04), and `all` also includes the gold-standard items (SUS and UEQ-S).

```{r show-params, echo=FALSE}
cat("Items used in this run (params$scales):", params$scales)
```

```{r load-data}

df <- read.csv(
  input_file,
  stringsAsFactors = FALSE,
  na.strings = c("NA", ""),
  check.names = FALSE
)

uxvis_groups <- list(
  Cognitive_value = c("P064", "P301", "P060", "P027"),
  Intuitiveness   = c("P088", "P026", "P4026A"),
  Visualization   = c("P106", "P089", "P098", "P097", "P022A", "P067")
)
gold_standard_groups <- list(
  SUS             = c("SUS_1", "SUS_2", "SUS_3", "SUS_4", "SUS_5",
                       "SUS_6", "SUS_7", "SUS_8", "SUS_9", "SUS_10"),
  UEQ_pragmatic   = c("UEQ_1", "UEQ_2", "UEQ_3", "UEQ_4"),
  UEQ_hedonic     = c("UEQ_5", "UEQ_6", "UEQ_7", "UEQ_8")
)

item_groups <- if (params$scales == "all") c(gold_standard_groups, uxvis_groups) else uxvis_groups

item_to_group <- data.frame(
  item  = unlist(item_groups, use.names = FALSE),
  group = rep(names(item_groups), lengths(item_groups)),
  stringsAsFactors = FALSE
)

all_items <- item_to_group$item

all_colors <- c(SUS = "#5E4EA3", UEQ_pragmatic = "#379EB8", UEQ_hedonic = "#4AAF77",
                Cognitive_value = "#E02884", Intuitiveness = "#F06128", Visualization = "#E2A600")
group_colors <- all_colors[names(item_groups)]
```

## Compute polychoric correlations (pairwise deletion)

`hetcor(..., use = "pairwise.complete.obs")` estimates each pairwise correlation from only
the rows where *that specific pair* of items is non-missing (pairwise deletion), rather than
dropping a participant from the whole matrix because a single item elsewhere is missing.
Every item is coerced to an ordered factor first, so every pair is estimated as a genuine
polychoric correlation (not Pearson or polyserial).

`std.err` is left off: computing polychoric standard errors requires a numerical Hessian,
and its finite-difference probing routinely pushes the bivariate-normal CDF differences into
underflow, producing `"NaNs produced"` warnings across most pairs. The `r` estimates
themselves are unaffected — only that Hessian step is unstable. Significance is instead
approximated below with the same t-approximation used for Pearson correlations
(`t = r * sqrt(n-2) / sqrt(1-r^2)`, `df = n-2`); this is a standard practical stand-in
for polychoric significance testing, not an exact analytic result.

`SUS_2, SUS_4, SUS_6, SUS_8, SUS_10` are the negatively-worded SUS items (Brooke, 1996) — on the
1–5 raw scale, a *higher* raw score means *worse* usability for these five, opposite every
other item in the matrix. They're reverse-coded (`6 - score`) before scoring, so every item
in the resulting matrix points the same "more positive" direction and signs are directly
comparable across the whole matrix (matches the reverse-coding already applied to the SUS
composite in the Pearson notebook).

```{r hetcor}
reverse_items <- c("SUS_2", "SUS_4", "SUS_6", "SUS_8", "SUS_10")

# 1. Select only targeting columns
item_df <- df[, all_items]

# 2. Invert SUS negative items (1-5 -> 6 - value)
# 
for (col in reverse_items) {
  if (col %in% colnames(item_df)) {
    item_df[[col]] <- 6 - as.numeric(item_df[[col]])
  }
}

# 3. Convert all columns to ordered factors 
item_df <- as.data.frame(lapply(item_df, ordered))

# 4. Executa a correlação policórica
het <- polycor::hetcor(item_df, use = "pairwise.complete.obs", std.err = FALSE, ML = FALSE)

```

```{r extract}
#r_mat <- het$correlations[all_items, all_items]  # enforce construct-block order
#n_mat <- het$n[all_items, all_items]

# Compute the correct pairwise sample size matrix (N)
#valid_data <- !is.na(item_df[, all_items])
#n_mat <- t(valid_data) %*% valid_data
#dimnames(n_mat) <- dimnames(r_mat)


#t_mat <- r_mat * sqrt(n_mat - 2) / sqrt(1 - r_mat^2)


#p_raw <- 2 * pt(-abs(t_mat), df = n_mat - 2)
#p_mat <- matrix(p_raw, nrow = nrow(r_mat), ncol = ncol(r_mat), dimnames = dimnames(r_mat))
#diag(p_mat) <- NA  # self-correlation (r = 1) makes t = Inf/NaN; not a meaningful p-value anyway

#groups_vec <- item_to_group$group[match(colnames(r_mat), item_to_group$item)]


#axis_text_element <- function(colors, angle = 0, hjust = 0.5, vjust = 0.5) {
#  element_text(angle = angle, hjust = hjust, vjust = vjust, colour = colors, face = "bold")
#}


r_mat <- het$correlations[all_items, all_items]  # enforce construct-block order

# Compute the correct pairwise sample size matrix (N)
valid_data <- as.matrix(!is.na(item_df[, all_items])) * 1
n_mat <- t(valid_data) %*% valid_data
dimnames(n_mat) <- dimnames(r_mat)

# Approximated t-test and p-values
t_mat <- r_mat * sqrt(n_mat - 2) / sqrt(1 - r_mat^2)
p_raw <- 2 * pt(-abs(t_mat), df = n_mat - 2)
p_mat <- matrix(p_raw, nrow = nrow(r_mat), ncol = ncol(r_mat), dimnames = dimnames(r_mat))
diag(p_mat) <- NA

groups_vec <- item_to_group$group[match(colnames(r_mat), item_to_group$item)]

# Theme helper for axis labels
axis_text_element <- function(colors, angle = 0, hjust = 0.5, vjust = 0.5, size = 12) {
  ggplot2::element_text(angle = angle, hjust = hjust, vjust = vjust, colour = colors, face = "bold", size = size)
}





```

## Results — long table

```{r long-table, results = 'asis'}
long <- as.data.frame(as.table(r_mat)) %>%
  rename(item1 = Var1, item2 = Var2, r = Freq) %>%
  filter(as.character(item1) < as.character(item2)) %>%
  mutate(
    group1 = item_to_group$group[match(as.character(item1), item_to_group$item)],
    group2 = item_to_group$group[match(as.character(item2), item_to_group$item)],
    within_construct = group1 == group2
  )

# Vectorized lookup via an integer index matrix (guaranteed to return a plain
# numeric vector, unlike mapply() over factor columns, which can degrade to a
# list column that write.csv() then can't serialize).
i_idx <- match(as.character(long$item1), colnames(p_mat))
j_idx <- match(as.character(long$item2), colnames(p_mat))
long$p <- p_mat[cbind(i_idx, j_idx)]
long$n <- n_mat[cbind(i_idx, j_idx)]

long <- long %>% arrange(group1, group2, item1, item2)

kable(long, digits = c(NA, NA, 3, NA, NA, NA, 4, 0))
```

## Heatmap — construct-block order

```{r heatmap, fig.width = 10, fig.height = 9}
plot_df <- as.data.frame(as.table(r_mat)) %>%
  rename(item1 = Var1, item2 = Var2, r = Freq) %>%
  mutate(item1 = factor(item1, levels = all_items),
         item2 = factor(item2, levels = rev(all_items)))

# axis label colour per item, in the same order as the axis ticks (x = all_items,
# y = rev(all_items))
item_colors_construct <- group_colors[groups_vec]

# a real "construct" legend requires a real aesthetic mapping — this dummy layer
# is invisible (alpha = 0) but its legend key is forced visible via override.aes,
# giving a colour-swatch legend for the axis label colours
construct_legend_df <- data.frame(
  item1 = factor(all_items[1], levels = levels(plot_df$item1)),
  item2 = factor(all_items[1], levels = levels(plot_df$item2)),
  construct = factor(names(group_colors), levels = names(group_colors))
)

p_heat_construct <- ggplot(plot_df, aes(item1, item2, fill = r)) +
  geom_tile(color = "grey80") +
  #geom_text(aes(label = sprintf("%.2f", r)), size = 1.6, colour = "grey20") +
  scale_fill_gradient2(
    low = "red", mid = "white", high = "blue",
    midpoint = 0, limits = c(-1, 1), na.value = "grey95"
  ) +
  geom_point(data = construct_legend_df, aes(x = item1, y = item2, colour = construct),
             alpha = 0, inherit.aes = FALSE) +
  scale_colour_manual(name = "Construct", values = group_colors) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4, shape = 15))) +
  coord_fixed() +
  theme_minimal() +
  theme(
    axis.text.x = axis_text_element(item_colors_construct, angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = axis_text_element(rev(item_colors_construct), size = 12),
    
    legend.text  = element_text(size = 12),     
    legend.title = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Polychoric correlations, ordered by construct",
    x = NULL, y = NULL, fill = "r"
  )

p_heat_construct


# --- save PNG to include in paper ---
ggsave(
  filename = file.path(output_dir, "heatmap_construct.png"),
  plot = p_heat_construct,
  width = 10,
  height = 9,
  dpi = 300
)



```

## Hierarchical clustering of items

Distance is `1 - r`: items that move together are placed close together. Now that the
negatively-worded SUS items are reverse-coded, `r` is directly comparable in sign across the
whole matrix, so the plain (non-absolute) distance is the right one to use. This is agnostic
to construct labels; it is purely data-driven, then compared against the intended construct
assignment.

```{r clustering}
dist_mat <- as.dist(1 - r_mat)
hc <- hclust(dist_mat, method = "average")

k <- length(item_groups)
empirical_cluster <- cutree(hc, k = k)

cat("Empirical cluster (k =", k, ") vs. intended construct:\n")
table(cluster = empirical_cluster, construct = groups_vec)
```

```{r dendrogram, fig.width = 11, fig.height = 6}
plot_dendrogram <- function() {
  if (requireNamespace("dendextend", quietly = TRUE)) {
    dend <- as.dendrogram(hc)
    label_order <- labels(dend)
    dend <- dendextend::set(dend, "labels_col",
                             group_colors[groups_vec[match(label_order, colnames(r_mat))]])
    dend <- dendextend::set(dend, "labels_cex", 0.8)
    plot(dend, main = "Hierarchical clustering of items (distance = 1 - r)")
    legend("topright", legend = names(group_colors), fill = group_colors,
           cex = 0.7, bty = "n")
  } else {
    plot(hc, hang = -1, xlab = "", sub = "",
         main = "Hierarchical clustering of items (distance = 1 - r)")
    mtext("install.packages('dendextend') for construct-colored labels", side = 1, cex = 0.7)
  }
}
plot_dendrogram()


# --- PNG (ideal to include in docs ---
png(file.path(output_dir, "dendrogram.png"), width = 10, height = 8, units = "in", res = 300)
plot_dendrogram()
dev.off()



```

## Heatmap — reordered by empirical clustering

Same matrix, same values — just reordered by `hc$order` instead of the intended construct
grouping, so you can visually compare the two structures.

```{r heatmap-clustered, fig.width = 10, fig.height = 9}
cluster_order <- colnames(r_mat)[hc$order]

plot_df_cluster <- as.data.frame(as.table(r_mat)) %>%
  rename(item1 = Var1, item2 = Var2, r = Freq) %>%
  mutate(item1 = factor(item1, levels = cluster_order),
         item2 = factor(item2, levels = rev(cluster_order)))

item_colors_cluster <- group_colors[groups_vec[match(cluster_order, colnames(r_mat))]]

construct_legend_df_cluster <- data.frame(
  item1 = factor(cluster_order[1], levels = levels(plot_df_cluster$item1)),
  item2 = factor(cluster_order[1], levels = levels(plot_df_cluster$item2)),
  construct = factor(names(group_colors), levels = names(group_colors))
)

p_heat_cluster <- ggplot(plot_df_cluster, aes(item1, item2, fill = r)) +
  geom_tile(color = "grey80") +
  geom_text(aes(label = sprintf("%.2f", r)), size = 1.6, colour = "grey20") +
  scale_fill_gradient2(
    low = "red", mid = "white", high = "blue",
    midpoint = 0, limits = c(-1, 1), na.value = "grey95"
  ) +
  geom_point(data = construct_legend_df_cluster, aes(x = item1, y = item2, colour = construct),
             alpha = 0, inherit.aes = FALSE) +
  scale_colour_manual(name = "Construct", values = group_colors) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4, shape = 15))) +
  coord_fixed() +
  theme_minimal() +
  theme(
    axis.text.x = axis_text_element(item_colors_cluster, angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = axis_text_element(rev(item_colors_cluster))
  ) +
  labs(
    title = "Polychoric correlations, reordered by empirical clustering",
    x = NULL, y = NULL, fill = "r"
  )

p_heat_cluster

# --- save PNG to include in paper ---
ggsave(
  filename = file.path(output_dir, "heatmap_empirical_cluster.png"),
  plot = p_heat_cluster,
  width = 10,
  height = 9,
  dpi = 300
)


```

## Save outputs (CSV + PDF)

A raw 31×31 numeric table isn't legible as a printed page, so the PDF renders the two
heatmaps and the dendrogram (one per page); the exact numbers are in the CSVs alongside it.

```{r save}
write.csv(r_mat, file.path(output_dir, "polychoric_correlation_matrix_r.csv"))
write.csv(p_mat, file.path(output_dir, "polychoric_correlation_matrix_p.csv"))
write.csv(n_mat, file.path(output_dir, "polychoric_correlation_matrix_n.csv"))
write.csv(long, file.path(output_dir, "polychoric_correlation_long.csv"), row.names = FALSE)





pdf(file.path(output_dir, "polychoric_correlation_matrices.pdf"), width = 10, height = 9)
print(p_heat_construct)
print(p_heat_cluster)
plot_dendrogram()
invisible(dev.off())
```

Note: this environment has no R installation to execute against, so this file has not been
run end-to-end — knit it locally (`rmarkdown::render(...)` or the Knit button in RStudio) and
flag anything that errors. `dendextend` (`install.packages("dendextend")`) is optional, only
for colored dendrogram labels, and degrades gracefully if missing. The two heatmaps' coloured
axis labels use `element_text(colour = <vector>)`, which triggers a harmless "Vectorized
input to element_text() is not officially supported" warning on knit — that's expected and
can be ignored; it was chosen over `ggtext::element_markdown()` after the latter proved
unreliable (intermittent literal `<span>` text, and a genuine `gridtext` error on repeat
runs) in testing.
